---
title: "Reading, Selected Hobo Temperature Logger"
author: "Chuck Burks"
date: "2023-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(readxl)
library(janitor)
theme_csb_halfwidth1 <- function(){
  theme_bw() +
    theme(axis.text.x = element_text(color = "black", size = 10),
          axis.text.y = element_text(color = "black", size = 10),
          axis.title.x = element_text(color = "black", size = 10),
          axis.title.y = element_text(color = "black", size = 10),
          legend.title = element_text(color = "black", size = 10),
          legend.text = element_text(color = "black", size = 9)
    )
}
```

## Read in Data

This document serves as a readable auto-reminder about how to fish necessary
information out of a Hobotemp data file. Here we are examining a file from 
a walk-in environmental chamber that started to have worrisome heating issues.

It is necessary to play with the directories a bit to get the correct file and 
file location. That can be done using autocomplete features in R/RStudio. Most 
of the needed functions are loaded with the tidyverse mega-library. The
clean_names() function of the janitor packages is also used to deal with messy
default variable names.


```{r }
getwd()
list.dirs()
```

```{r }
dat_in <- read_csv("Queue 1 2023-07-19 06_13_48 PDT (Data PDT).csv")
dat_in
```

This throws a warning, but it works. Additional clean-up steps are applied

```{r }
dat_in <- clean_names(dat_in)
dat_in
```

```{r }
dat_in$date_time_pst_pdt <- as.POSIXct(mdy_hms(dat_in$date_time_pst_pdt))

dat_in <- dat_in %>% 
  rename(deg_c = 3,
         rh = ch_2_rh_percent)

dat_in
```

Also examining the end of the file

```{r, echo=FALSE }
tail(dat_in)
```

## Plot tempertures over the entire time frame

For Rmarkdown the graphic is not saved, but the code in the accompanying 
R script includes formatting suitable for a journal 1 column (half page width)
format, in part because I think that also sizes well if sent on a cell phone.

```{r}
ggplot(dat_in, aes(x = date_time_pst_pdt, y = deg_c)) +
  geom_line() +
  labs(x = "",
       y = "Degree Celcius",
       caption = "Queue2 in H123") +
  theme_csb_halfwidth1()
```
This output indicates that the temperature held steady for weeks, and started 
to spike more recently. I want to expand the X axis only within this
time frame. I will use therefore use lubridate functions (loads with tidyverse)
like month(), epiweek(), wday(), etc. to narrow the data set to times of 
interest. In this case we start by examining the last few epiweeks (weeks of 
the year, see ?epiweek)

```{r}
dat_in %>% 
  mutate(wk = epiweek(date_time_pst_pdt)) %>% 
  group_by(wk) %>% 
  summarise(first_reading = min(date_time_pst_pdt)) %>% 
  tail(.)
```

The data end with epiweek 17, and that captures the last 5 or so days of data.

```{r}
dat_in %>% 
  filter(epiweek(date_time_pst_pdt) == 17) %>% 
  ggplot(., aes(x = date_time_pst_pdt, y = deg_c)) +
  geom_line() +
  labs(x = "",
       y = "Degree Celcius",
       caption = "Queue2 in H123") +
  theme_csb_halfwidth1()
```
So I have had the spiking problem for at least a week, but not a month. 
I can save these charts, pass them to somewhere that I can get them into the 
camera roll on my phone, and that way share them with maintenance and 
contractors in a quickly readable format. See "03script4-read-hobo-output-csv.R
for the ggsave code.