# Muliple Data Frames To and From Multipe Sheets in an Excel Document

This is intended as a vignette showing how to read in multiple tabs from an 
Excel document, then recombine them into a single data frame. 

```{r include=FALSE}
library(openxlsx)
library(dplyr)
library(readxl)
library(purrr)
```

## 1. Load and characterize internal iris data set 
```{r}
irises <- iris
head(irises)
```

Remind ourselves of the species
```{r}
unique(irises$Species)
```

## 2. Divide Iris into separate data frames, one per species 

See https://dplyr.tidyverse.org/reference/group_split.html

Using base::split() creates a list containing three named data frames
```{r}
y <- split(irises, irises$Species)
names(y)
head(y$setosa)
```

dplyr::group_split() is another option, but the data frame elements  are not
named. They are just mapped with subscripts. Therefore split() seems to serve the present 
purpose better
```{r}
x <- irises %>% 
  group_split(Species)
```

## 3. Export the data frames to sheets in a single spreadsheet doc

Working from a list of data frames, this is one short and simple function. See
https://www.r-bloggers.com/2022/02/export-data-frames-into-multiple-excel-sheets-in-r/

```{r}
openxlsx::write.xlsx(y, file = "data.xlsx")

```

## 4. Read back the 3 data frames from 3 sheets in a single spreadsheet

This r-blogger entry addresses how to approach this with lapply or with 
purr::map(). See:

https://www.r-bloggers.com/2022/07/read-data-from-multiple-excel-sheets-and-convert-them-to-individual-data-frames/

Read Excel sheets into a list of three data frames using lapply
```{r}
df_list1 <- lapply(readxl::excel_sheets("data.xlsx"), function(x)
  read_excel("data.xlsx", sheet = x)
)

df_list1$setosa
```

Read Excel sheets into a list of three data frames using purrr::map()
```{r}
df_list2 <- map(set_names(excel_sheets("data.xlsx")),
               read_excel, path = "data.xlsx"
)
df_list2$setosa
```

## 5. Join the data frames end-to-end

Join end-to-end using base::do.call. Creates a 150-row data frame from the 
list of three 50-row data frames.
```{r}
irises2 <- do.call(rbind, df_list2)
irises2
```

Join end-to-end using dplyr::bind_rows. This accomplishes the same thing, but
dplyr::bind_rows() has other advantages over do.call(). See:
 https://dplyr.tidyverse.org/reference/bind.html
```{r}
irises3 <- bind_rows(df_list2)

```

For greater depth, see also:
https://www.r-bloggers.com/2021/08/r-a-combined-usage-of-split-lapply-and-do-call/

## 6. Convert the list of 3 data frames to 3 separate data frames

Convert the list to three separate data frames. See:
See https://stackoverflow.com/questions/13795526/return-elements-of-list-as-independent-objects-in-global-environment

The list generated with lapply() does not work with this method
```{r}
# list2env(df_list1,globalenv())
```

The error message generated is: "Error in list2env(df_list1, globalenv()) : 
names(x) must be a character vector of the same length as x"

The list generated with purrr::map does work
```{r}
list2env(df_list2,globalenv())
```



