---
title: 'How to save (and load) datasets in R: an overview'
author: "Sascha Wolfer"
date: "2023-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Primary reference
http://rcrastinate.rbind.io/post/how-to-save-and-load-data-in-r-an-overview/

### See also
https://statisticsglobe.com/r-save-load-rdata-workspace-file
https://youtu.be/svgEpRzhG7M

## What I will show you

In this post, I want to show you a few ways how you can save your datasets in R. 
Maybe, this seems like a dumb question to you. But after giving quite a few R 
courses mainly - but not only - for R beginners, I came to acknowledge that the 
answer to this question is not obvious and the different possibilites can be 
confusing. In this post, I want to give an overview over the different 
alternatives and also state my opinion which way is the best in which situation.

## What are we going to do?

I will show you the following ways of saving or exporting your data from R:
 - Saving it as an R object with the functions save() and saveRDS()
 - Saving it as a CSV file with write.table() or fwrite()
 - Exporting it to an Excel file with WriteXLS()
 
For me, these options cover at least 90% of the stuff I have to do at work. 
So I hope that it’ll work for you, too.

## Preparation: Load some data

I will use some fairly (but not very) large dataset from the car package. The 
dataset is called MplsStops and holds information about stops made by the 
Minneapolis Police Department in 2017. Of course, you can access this dataset 
by installing and loading the car package and typing MplsStops. However, I want 
to simulate a more typical workflow here. Namely, loading a dataset from your 
disk (I will load it over the WWW). The dataset is also available from GitHub:  


```{r}
data <- read.table("https://vincentarelbundock.github.io/Rdatasets/csv/carData/MplsStops.csv",
                   sep = ",", header = T,
                   row.names = 1)

kableExtra::scroll_box(knitr::kable(head(data), row.names = F),
           width = "100%", height = "300px")
```

We now have a dataset with over 50,000 rows (you can scroll through the 
first 6 of them in the box above) and 14 variables in our global environment 
(the ‘workspace’). ust for the sake of simulating a real workflow, I will do 
some very light data manipulation. Here, I’m assigning a new column 
data$gender.not.known which is TRUE whenever data$gender is "Unknown" or NA.

```{r}
data$gender.not.known <- is.na(data$gender) | data$gender == "Unknown"
```

As I wrote above: Saving the current state of your dataset in R makes sense 
when all the preparations take a lot of time. If they don’t, you can just run 
your pre-processing code every time you are getting back to analyzing the 
dataset. In the scope of this post, let’s suppose that the calculation above 
took veeeery long and you absolutely don’t want to run it everytime.

## Option 1: Save as an R object

Whenever I’m the only one working on a project or everybody else is also 
using R, I like to save my datasets as R objects. Basically, it’s just saving 
a variable/object (or several of them) in a file on your disk. There are two 
ways of doing this:
 1. Use the function save() to create an .Rdata file. In these files, you can store several variables.
 2. Use the function saveRDS() to create an .Rds file. You can only store one variable in it.

### Option 1.1: save()

# You can save your data simply by doing the following:

```{r}
save(data, file = "data.Rdata")
```

By default, the parameter compress of the save() function is turned on. That 
means that the resulting file will use less space on your disk. However, if it 
is a really huge dataset, it could take longer to load it later because R first 
has to extract the file again. So, if you want to save space, then leave it as 
it is. If you want to save time, add a parameter compress = F.# 

(First remove the file from the Global Environment to demonstrate that you
are, in fact, reloading it)

```{r}
rm(data)
```

If you want to load such an .Rdata file into your environment, simply do

```{r}
load(file = "data.Rdata")
```

Then, the object is available in your workspace with its old name. Here, the 
new variable will also have the name data. With save() You can also save 
several objects in one file. Let’s duplicate data to simulate this.

```{r}
data2 <- data

save(list = c("data","data2"), file = "data.Rdata")
```

Now, if you do load("data.Rdata"), you will have two more objects in your 
workspace, namely data and data2.

```{r}
rm(data,data2)

load(file = "data.Rdata")
```

### Option 1.2: saveRDS()

This is the second option of saving R objects. saveRDS() can only be used to 
save one object in one file. The “loading function” for saveRDS() is readRDS(). 
Let’s try it out.

```{r}
saveRDS(data, file = "data.Rds")
data.copy <- readRDS(file = "data.Rds")
```

### The difference between save() and saveRDS()

So, you might ask “why should I use saveRDS() instead of save()”? Actually, I 
like saveRDS() better - for one specific reason that you might not have noticed 
in the calls above. When we use load(), we do not assign the result of the 
loading process to a variable because the original names of the objects are 
used. But this also means that you have to “remember” the names of the 
previously used objects when using load().

When we use readRDS(), we have to assign the result of the reading process to a 
variable. This might mean more typing but it also has the advantage that you 
can choose a new name for the variable to integrate it in into the rest of the 
new script more smoothly. Also, it is more similar to the behavior of all the 
other “reading functions” like read.table(): for these, you also have to assign 
the result to a variable. The only advantage of save() really is that you can 
save several objects into one file - but in the end it might be better to have 
one file for one object. This might be more clearly organized.

(The rest of the blog entry is about saving csv and excel files. I tend to 
use write.csv for the former. I do not use the WriteXLS package as often 
because I prefer to output to csv, but I have used WriteLXS and it is good
to be aware of.)

