---
title: "Transform Dataframe to Markdown Table"
output: html_document
date: "2023-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function to make a dataframe into a Markdown plot

This is a direct appropriation of a blog post by Tomaz Kastrun. See:

https://tomaztsql.wordpress.com/2023/03/26/little-useless-useful-r-functions-transforming-dataframe-to-markdown-table/ 

The purpose is to transform a dataframe into a markdown table, which would then be displayed as a regular table if the markdown document is viewed in formatted form.

The function is as follows:

```{r}
df_2_MD <- function(your_df){
  cn <- as.character(names(your_df))
  headr <- paste0(c("", cn),  sep = "|", collapse='')
  sepr <- paste0(c('|', rep(paste0(c(rep('-',3), "|"), collapse=''),length(cn))), collapse ='')
  st <- "|"
    for (i in 1:nrow(your_df)){
      for(j in 1:ncol(your_df)){
        if (j%%ncol(your_df) == 0) {
          st <- paste0(st, as.character(your_df[i,j]), "|", "\n", "" , "|", collapse = '')
        } else {
        st <- paste0(st, as.character(your_df[i,j]), "|", collapse = '')
        }
      }
    }
  fin <- paste0(c(headr, sepr, substr(st,1,nchar(st)-1)), collapse="\n")
  cat(fin)
}  
```


## Demostration of the function

Get a small part of the iris data set for demonstration.

```{r}
short_iris <- iris[1:3,1:5]
short_iris
```

Apply the function
```{r iris}
df_2_MD(short_iris)
```
For a demonstration of the utility of this output, see

https://github.com/ChuckBV/Work_misc 